---
title: "Assignement - Bayesian Statistics"
author: "Matěj Tvarůžka"
date: "2026-02-16"
format: html
---

# Coexistence (syntopy) of sister species on a local scale

**Question: How can two closely sister species of birds co-occurr at the same site?**

::: callout-note
## Steps

1.  Description of data and problem

2.  Basic visualization

3.  Model evaluation
:::

## 1. Data and problem

To estimate the importance of preditors for syntopy, we take a look at the probabilities of species co-occurence at the same sites. We need to select only the sites within the range overlap (sympatry) of these two species (otherwise the results will be biased -see Remeš & Harmáčková 2023).

We can imagine the range overlap being the bag, from which we randomly draw co-occurence states. There are four possible outcomes of that drowing: 1) both species are present, 2) species A is present, 3) species B is present or 4) both species are missing at the site. The most accurate probability distribution for this process is the Fisher's non-central hypergeometric distribution, where *N* stands for the population size (number of sites within the sympatric range), *K* is the number of co-occurrence states (*K* = 4), *n* is the number of draws and *k* stands for number of co-occurrences.

$$
\begin{aligned}
K_{ij} &\sim \text{fnchypg}(\vec{m}_i, N - \vec{m}_i, \vec{m}_j, e^\alpha) \\
\alpha &= \beta_0 + \lambda_i + \lambda_j + \beta X \\
\beta_k &\sim \text{N}(0, 1) \\
\beta_0 &\sim \text{N}(0, 4) \\
\lambda_i &\sim \text{N}(0, 1)
\end{aligned}
$$\

```{r}
library(rethinking)
```

```{r}

```
